---
title: "May31"
author: "Julin N Maloof"
date: "5/28/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises 10.5

### 1
_How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame)._

```{r}
library(tidyverse)
mtcars
as_tibble(mtcars)
class(mtcars)
class(as_tibble(mtcars))
```

In the Rstudio Rmd interacre you can;t tell just by printing it (other than the row names go away).  At the console it is clear because we are told that a tiblle is a tibble.  Otherwise use `class()`


### 2
_Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?_

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

```{r}
dft <- as_tibble(df)
dft$x
dft[, "xyz"]
dft[, c("abc", "xyz")]
```

tibles require exact matches for column names.

tible extactions always always return a tibble, whereas dataframes can return a vector.

### 3
_If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?_

```{r}
var <- "hwy"
data(mpg)
mpg
get(var,mpg)
select(mpg,matches(var))
subset(mpg,select=var)
mpg[var]
mpg[[var]]
mpg[,var]
```


### 4
_Practice referring to non-syntactic names in the following data frame by:_

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

_1. Extracting the variable called 1._

```{r}
annoying$`1`
annoying[,"1"]
annoying %>% select(`1`)
```

_2. Plotting a scatterplot of 1 vs 2._
```{r}
qplot(data=annoying,x=`1`,y=`2`)
```


_3. Creating a new column called 3 which is 2 divided by 1._

```{r}
annoying2 <- annoying %>% mutate(`3` = `2`/`1`) 
annoying2
```


_4. Renaming the columns to one, two and three._

```{r}
lessAnnoying <- annoying2 %>% rename(one=`1`, two=`2`, three=`3`)
lessAnnoying
```


### 5
_What does tibble::enframe() do? When might you use it?_

converts a vector to a tibble

```{r}
myVector <- c(a=1,b=2,c=3)
myVector
myVector %>% enframe()
```


### 6
_What option controls how many additional column names are printed at the footer of a tibble?_

`tibble.max_extra_cols`


```{r}
library(nycflights13)
flights
options(tibble.max_extra_cols=3)
flights
```

