---
title: "August 23rd"
author: "Julin N Maloof"
date: "8/22/2017"
output: 
  html_document: 
    keep_md: yes
---

## 19.5.5 exercises

### 1
_What does commas(letters, collapse = "-") do? Why?_

This will return `a-b-c-d-e` etc because we are over-riding the default

### 2
_It’d be nice if you could supply multiple characters to the pad argument, e.g. rule("Title", pad = "-+"). Why doesn’t this currently work? How could you fix it?_

```{r}
rule <- function(..., pad = "-") {
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep = "")
}

rule("Title", pad = "-+")
```

Fails because the width of the pad is not being taken into account

```{r}
rule <- function(..., pad = "-") {
  title <- paste0(...)
  width <- (getOption("width") - nchar(title) - 5)/nchar(pad)
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep = "")
}

rule("Title", pad = "-+")
```

### 3
_What does the trim argument to mean() do? When might you use it?_
```{r}
?mean
```

Trims the fraction of observations from each end of the input.  Would be good if input was multiple observations from a device where the beginning and end points were noisy.

### 4
_The default value for the method argument to cor() is c("pearson", "kendall", "spearman"). What does that mean? What value is used by default?_

This is providing the various options and the first one is used by default.

See 
```{r}
?match.arg
```
which cor() uses

