---
title: "Untitled"
author: "Julin N Maloof"
date: "7/17/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringr)
```

## 14.4.2 Exercises

### 1
_Find all words that start or end with x_
```{r}
mywords <- c(words,"xylophone","xeriscape","xanax")
str_subset(mywords,"^x|x$")

mywords[str_detect(mywords,"^x") | str_detect(mywords,"x$")]
```

_Find all words that start with a vowel and end with a consonant._

```{r}
str_subset(mywords,"^[aeiou].?[^aeiou]$")

mywords %>% extract(str_detect(.,"^[aeiou]") & ! str_detect(.,"[aeiou]$"))
```

_Are there any words that contain at least one of each different vowel?_

Not sure how to do this in single regex

```{r}
mywords %>% extract(.,
                    str_detect(.,"a") & 
                      str_detect(.,"e") & 
                      str_detect(.,"i") & 
                      str_detect(.,"o") & 
                      str_detect(.,"u") 
)
    
```


### 2

_What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)_

```{r}
words.df <- tibble(
  word=mywords)

#total vowels
words.df %>% mutate(vowels=str_count(word,"[aeiou]")) %>% arrange(desc(vowels))

#average vowels
words.df %>% 
  mutate(vowels=str_count(word,"[aeiou]"),
         letters=str_length(word),
         proportion=vowels/letters) %>% 
  arrange(desc(proportion))
```

## 14.4.3.1 Exercises

### 1 
_In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem._
```{r}
colors <- c("red", "orange", "yellow", "green", "blue", "purple")
color_match <- str_c("\\b",colors,"\\b", collapse = "|")
color_match
```

```{r}
more <- sentences[str_count(sentences, color_match) > 1]
str_view_all(more, color_match)
```

### 2
_From the Harvard sentences data, extract:_

_The first word from each sentence._
_All words ending in ing._
_All plurals._

```{r}
str_extract(sentences,"^\\w+\\b")
```

```{r}
ing <- "\\b\\w*ing\\b"
sentences %>%
  str_subset(ing) %>%
  str_extract_all(ing) %>% 
  unlist() %>%
  unique()
```

```{r}
plurals <- "\\b\\w*[^aeiousy' ]s\\b"
sentences %>% 
  str_subset(plurals) %>% #only those sentences with a plural
  str_extract_all(plurals,simplify=FALSE) %>%
  unlist() %>%
  unique() %>%
  .[!str_detect(.,"help|its")]
```

## 14.4.1 Exercises

### 1 
_Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word._

```{r}
numbers <- str_replace_all("(one two three four five six seven eight nine ten eleven twelve thirteen fourteen fifteen sixteen seventeen eighteen nineteen twenty)"," ","|")
numbers

pattern <- str_c(numbers,"\\s(\\w+)")
pattern

sentences %>% 
  str_subset(pattern) %>%
  str_match(pattern)
```

### 2
_Find all contractions. Separate out the pieces before and after the apostrophe._
```{r}
pattern0 <- "([Ii]t's|[Ll]et's)|((\\w+)'([^s ]+))" #match all words with a "'" except for possesives
pattern <- "(\\w+)'(\\w+)"

sentences %>% 
  str_subset(pattern0) %>%
  str_match(pattern)
```

