---
title: "Untitled"
author: "Julin N Maloof"
date: "5/16/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r cars}
library(nycflights13)
library(tidyverse)
```

## Scratchpad

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)

by_dest

delay <- filter(delay, count > 20, dest != "HNL")

delay
```

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")

```

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

```{r}
not_cancelled %>% 
  count(dest)
```

```{r}
not_cancelled %>%
  group_by(flight) %>%
  mutate(
    late_15 = mean(arr_delay == 15),
    early_15 = mean(arr_delay == -15)
    ) %>%
  filter(late_15 == .5, !duplicated(flight)) %>%
  select(carrier, flight, origin, dest) # because the other columns are specific to particular days
```

```{r}
not_cancelled %>%
  group_by(flight) %>%
  summarize(
    late_15 = mean(arr_delay == 15),
    early_15 = mean(arr_delay == -15)
    ) %>%
  filter(late_15 == .5)
```


## 5.6.7 Exercises

### 1

Skipping it

### 2

_Come up with another approach that will give you the same output as below without using `count()`_
```{r}
not_cancelled %>% 
  count(dest)
not_cancelled %>% 
  count(tailnum, wt = distance)
```

```{r}
not_cancelled %>% group_by(dest) %>% summarize(n=n())
not_cancelled %>% group_by(tailnum) %>% tally(wt=distance)

```

### 3

_Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?_

arr_delay is the more important column.  Maybe the departure delay wasn't recorded?

### 4

_Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?_

```{r}
cancelled_by_day <- flights %>% group_by(year,month,day) %>%
  summarize(cancelled=sum(is.na(dep_delay) | is.na(arr_delay)),
            avg_arr_delay=mean(arr_delay,na.rm=TRUE))
cancelled_by_day
qplot(x=avg_arr_delay,y=cancelled,data = cancelled_by_day) + geom_smooth()
```

### 5 

_Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about_ `flights %>% group_by(carrier, dest) %>% summarise(n()))`

```{r}
not_cancelled %>% group_by(carrier) %>% summarize(mean_delay=mean(arr_delay)) %>% arrange(desc(mean_delay))
```

```{r}
not_cancelled %>% group_by(carrier,dest) %>% summarize(mean_delay=mean(arr_delay)) %>% arrange(desc(mean_delay)) %>%
  ggplot(aes(x=carrier,y=dest,fill=mean_delay)) + geom_raster() + scale_fill_gradient()
```


### 6

_What does the sort argument to count() do. When might you use it?_

Sorts...(duh).  Use it when you want to sort the counts....
```{r}
not_cancelled %>% count(dest)

not_cancelled %>% count(dest,sort = TRUE)

not_cancelled %>% count(dest) %>% arrange(desc(n))
```

## 5.7.1 Exercises

### 1

Skipping

### 2

_Which plane (tailnum) has the worst on-time record?_


```{r}
not_cancelled %>% group_by(tailnum) %>%
  mutate(avg_arr_delay = mean(arr_delay)) %>% 
  arrange(desc(avg_arr_delay))
```

If we want to limit to tailnum with > 25 flights
```{r}
not_cancelled %>% group_by(tailnum) %>%
  mutate(avg_arr_delay = mean(arr_delay), n=n()) %>% 
  filter(n > 24, !duplicated(tailnum)) %>%
  arrange(desc(avg_arr_delay)) %>%
  select(tailnum,avg_arr_delay,n)
```

### 3
_What time of day should you fly if you want to avoid delays as much as possible?_

```{r}
pl <- ggplot(not_cancelled,aes(x=sched_dep_time,y=dep_delay)) + geom_smooth()
pl
```

```{r}
not_cancelled %>% mutate(dep_hour = sched_dep_time %/% 100) %>%
  group_by(dep_hour) %>%
  summarize(mean_arr_delay = mean(arr_delay)) %>%
  ggplot(aes(x=dep_hour,y=mean_arr_delay)) + geom_col()
```

### 4

_For each destination, compute the total minutes of delay. For each, flight, compute the proportion of the total delay for its destination._

```{r}
not_cancelled %>% group_by(dest) %>%
  mutate(total_delay = sum(arr_delay)) %>%
  group_by(flight) %>%
  mutate(proportion_delay = arr_delay / total_delay) %>%
  arrange(dest,desc(proportion_delay))
```

