---
title: "June 14 Tidy Data 12.1 - 12.4"
author: "Julin N Maloof"
date: "6/13/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## 12.2.1 Exercises

### 1.

### 2. 

## 12.3.1 Exercises

### 1 
_Why are gather() and spread() not perfectly symmetrical?
Carefully consider the following example:_

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks
stocks %>% 
  spread(year, return)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
```
`year` gets converted to a character (since it was cycled through a column name)

the order chagnes

the `convert` argument converts data types, so:

```{r}
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`, convert=TRUE)
```


### 2 

```{r, eval=FALSE}
table4a
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
#> Error in combine_vars(vars, ind_list): Position must be between 0 and n
```
This fails because the column names `1999` and `2000` need to be backticked.  R is interpreting them as numeric.

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
#> Error in combine_vars(vars, ind_list): Position must be between 0 and n
```

### 3
_Why does spreading this tibble fail? How could you add a new column to fix the problem?_

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
people
```

```{r, eval = FALSE}
people %>% spread(key = key,value = value)
```

Fails because "Phillip Woods" is represted twice

```{r}
people  %>% group_by(name,key) %>% 
  mutate(observation=row_number()) %>% 
  spread(key = key,value = value)
```

Not perfect because we don't know at which age the height was measured...

### 4 

_Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?_

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
preg
```

Male and female need to be gathered 

```{r}
preg %>% gather(male, female, key="gender", value=cases)
```

Potentially gathered and spread..

```{r}
preg %>% gather(male, female, key="gender", value=cases) %>%
  spread(key = pregnant, value=cases) %>%
  rename(not_pregnant=no, pregnant=yes)
```
```

## 12.4.3 Exercises

### 1.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
separate(x, c("one", "two", "three")) # g gets lost

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
separate(x, c("one", "two", "three"),extra="drop") # g gets lost, no warning

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
separate(x, c("one", "two", "three"), extra="merge") # f,g remain unseparated

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
separate(x, c("one", "two", "three")) #NA for row 2, column 3 + warning

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
separate(x, c("one", "two", "three"),fill="right") #NA for row 2, column 3 no warning 

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
separate(x, c("one", "two", "three"),fill="left") #NA for row 2, column 1 no warning 

```


### 2.

Keep the original data column intact

### 3

Separate separates by definine the separator, extract separates by regular rexpression groups.

Well because unite you always want to define the separator...

