---
title: "May 03"
author: "Julin N Maloof"
date: "5/3/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3.6.1 Exercises

1.

```{r}
library(ggplot2)
library(tidyverse)
?geom_line
?geom_boxplot
?geom_histogram
?geom_area
```

2.

points and line colored by drv, showing relationship between displacement and highway MPG

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

3. it prevents the legend from being made:

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(se = FALSE,show.legend = FALSE)
```

4.
shows the standard error of the estimate via shading.

5.
will look the same.  Same aesthetics to both.

6.
```{r}
pl <- ggplot(mpg,aes(x=displ,y=hwy))
pl <- pl + geom_point()
pl <- pl + geom_smooth(se=FALSE)
pl
```

```{r}
pl <- ggplot(mpg,aes(x=displ,y=hwy,group=drv))
pl <- pl + geom_point()
pl <- pl + geom_smooth(se=FALSE)
pl
```

```{r}
pl <- ggplot(mpg,aes(x=displ,y=hwy,color=drv))
pl <- pl + geom_point()
pl <- pl + geom_smooth(se=FALSE)
pl
```

```{r}
pl <- ggplot(mpg,aes(x=displ,y=hwy))
pl <- pl + geom_point(aes(color=drv))
pl <- pl + geom_smooth(se=FALSE)
pl
```

```{r}
pl <- ggplot(mpg,aes(x=displ,y=hwy))
pl <- pl + geom_point(aes(color=drv))
pl <- pl + geom_smooth(aes(lty=drv),se=FALSE)
pl
```

```{r}
pl <- ggplot(mpg,aes(x=displ,y=hwy))
pl <- pl + geom_point(aes(fill=drv),color="white",stroke=1,shape=21,size=2)
pl
```

## 3.7.1 exercises

1. `geom_pointrange()`

2. geom_col is like geom_bar but it uses stat_identity as default.  The code below produces the same plots:

```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

```{r}
ggplot(data = demo) +
  geom_col(mapping = aes(x = cut, y = freq))
```


4.

stat_smooth computes y, ymin, ymax, and se.  It's behaviour is controlled by method, se, formula, n, span, level

5.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```

every category is its own group and proportions are calcualted relative to that.

## 3.8.1 Exercises

1.


```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

Points are overlapped.

Fix with

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

2. Jitter is controlled with the `width` and `height` arguments?geom

3,


```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
```

Count deals with overlapped data by binning them and adjusting point size.

4.

The default adjustment for box-plot is dodge:

```{r}
ggplot(mpg,aes(x=class,y=hwy,fill=drv)) + geom_boxplot()
```

## 3.9.1 Exercises

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

1. Turn a stacked bar chart into a polar chart

```{r}
pl <- ggplot(mpg,aes(x=as.factor(cyl), y = hwy, fill = drv))
pl + geom_col()
pl + geom_col() + coord_polar() + labs(x=NULL,y=NULL)
```
 
Not very sensical because mileage is getting summed...but the coverted graph looks cool

2. `labs()` sets the x and y axis labels (and also title and subtitle)

3. `coord_quickmap()` and `coord_map()` have the same goal, but `quickmap` is a fast approximation and works best for small areas near the equator

4. 

```{r}
pl <- ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline()
pl +  ggtitle("standard")
pl + coord_fixed() + ggtitle("coord_fixed")
pl + coord_fixed() + ggtitle("coord_fixed") + geom_smooth(method = "lm", se=FALSE)
```

There is a ~ linear relationship but it is offset from the diagonal and perhaps has a slope a bit higher than one.  

coord_fix insures that there is a 1:1 relationshp between the scales on the x and y axis.

geom_abline() adds a lins from point a to point b.  By default it is the diagonal.

## 4.4 Exercises

1.  There is a typo, "1" instead of "i" in "var1able"

2. 

```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```


3. keyboard shortcuts.  also found under help menu.