---
title: "Sept 13"
author: "Julin N Maloof"
date: "9/13/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

## 21.5.3 Exercises

### 1. _Write code that uses one of the map functions to:_

_Compute the mean of every column in mtcars._

```{r}
mtcars %>% map_dbl(mean)
```

_Determine the type of each column in nycflights13::flights._

```{r}
nycflights13::flights %>%
  map_chr(typeof)
```

_Compute the number of unique values in each column of iris._

```{r}
iris %>% map_int(~length(unique(.)))
```

_Generate 10 random normals for each of $mu$ = -10, 0, 10, and 100._

```{r}
mus <- c(-10,0,10,100)
mus %>% map(~rnorm(10,.)) 
#or
mus %>% map(rnorm,n=10)
```

### 2. _How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?_

```{r}
iris %>% map_lgl(is.factor)
```


### 3. _What happens when you use the map functions on vectors that aren’t lists? What does map(1:5, runif) do? Why?_

This works fine; map is applied to each element of an atomic list

```{r}
map(1:5,runif)
```

### 4. _What does map(-2:2, rnorm, n = 5) do? Why? What does map_dbl(-2:2, rnorm, n = 5) do? Why?_

```{r}
map(-2:2, rnorm, n = 5) 
```

so map will fill in the chunked data into the first empty argument?  cool!


```{r, error=TRUE}
map_dbl(-2:2, rnorm, n = 5)
```
error because output of each chunk is a vector of length 5

### 5 _Rewrite map(x, function(df) lm(mpg ~ wt, data = df)) to eliminate the anonymous function._

```{r}
mtcars %>%
  split(.$cyl) %>%
  map(~lm(mpg~wt, data = .))
```

## 21.9.3 Exercises

### 1.
_Implement your own version of every() using a for loop. Compare it with purrr::every(). What does purrr’s version do that your version doesn’t?_

```{r, error=TRUE}
my_every <- function(x,test) {
  results <- vector("logical",length=length(x))
  for(i in seq_along(x))  {
    results[[i]] <- test(x[[i]])
  }
  if(sum(results) == length(x)) TRUE else FALSE
}

x <- 1:5

my_every(x, is.integer)

my_every(iris, is.factor)

every(x, "> 5")

my_every(x, "> 5")

```

my_every can only take functions

But...

```{r}
my_every <- function(x,test) {
  results <- vector("logical",length=length(x))
  for(i in seq_along(x))  {
    if(is.function(test)) 
      results[[i]] <- test(x[[i]])
    else
      eval(parse(text=paste(x,test)))
  }
  if(sum(results) == length(x)) TRUE else FALSE
}

x <- 1:5

my_every(x, is.integer)

my_every(iris, is.factor)

every(x, "> 5")

my_every(x, "> 5")
```

### 2. _Create an enhanced col_sum() that applies a summary function to every numeric column in a data frame._

Didn't we already do this?

### 3._A possible base R equivalent of col_sum() is:_


```{r}
col_sum3 <- function(df, f) {
  is_num <- sapply(df, is.numeric)
  df_num <- df[, is_num]

  sapply(df_num, f)
}
```


But it has a number of bugs as illustrated with the following inputs:

```{r, error=TRUE}
df <- tibble(
  x = 1:3, 
  y = 3:1,
  z = c("a", "b", "c")
)
# OK
col_sum3(df, mean)
# Has problems: don't always return numeric vector
col_sum3(df[1:2], mean)
col_sum3(df[1], mean)
col_sum3(df[0], mean)
```

What causes the bugs?

This is because extracting from a data frame with [] changes types.
