---
title: "Sept-06"
author: "Julin N Maloof"
date: "9/5/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## 21.2.1

### 1

_Write for loops to:_

_1. Compute the mean of every column in mtcars._

```{r}
result <- vector("double",ncol(mtcars))
names(result) <- colnames(mtcars)
for(i in seq_along(mtcars)) {
  result[i] <- mean(mtcars[[i]],na.rm=TRUE)
}
result
```

_2. Determine the type of each column in nycflights13::flights._

```{r}
flights <- nycflights13::flights
result <- vector("character",ncol(flights))
names(result) <- colnames(flights)
for(i in seq_along(flights)) {
  result[i] <- typeof(flights[[i]])
}
result
```

_3. Compute the number of unique values in each column of iris._

```{r}
result <- vector("integer",ncol(iris))
names(result) <- colnames(iris)
for(i in seq_along(iris)) {
  result[i] <- length(unique(iris[[i]]))
}
result
```

_4. Generate 10 random normals for each of  `μ=−10 ,  0 ,  10 , and  100.`_

```{r}
means <- c(-10,0,10,100)
result <- matrix(NA,ncol=length(means),nrow=10)
for(i in seq_along(means)) {
  result[,i] <- rnorm(10,mean=means[i])
}
result
```

### 2
_Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:_
```{r}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}
out
```

```{r}
out <- stringr::str_c(letters,collapse="")
out
```

```{r}
x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))
sd
```

```{r}
sd <- sd(x)
sd
```

```{r}
x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
out
```

```{r}
cumsum(x)
```

### 3
_Combine your function writing and for loop skills:_

### 4
_It’s common to see for loops that don’t preallocate the output and instead increase the length of a vector at each step:_

```{r}
x <- lapply(floor(runif(10000,min = 1,max = 1000)), rnorm) # a list of vectors of variable lengths

system.time({
  output <- vector("integer", 0)
  for (i in seq_along(x)) {
    output <- c(output, length(x[[i]])) #changed from lengths to length because that made more sense to me
  }
})

head(output)

```

```{r}
system.time({
  output <- vector("integer", length(x))
  for (i in seq_along(x)) {
    output[i] <- length(x[[i]]) #changed from lengths to length because that made more sense to me
  }
})

head(output)
```


## 21.3.5

### 1
_Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, files <- dir("data/", pattern = "\\.csv$", full.names = TRUE), and now want to read each one with read_csv(). Write the for loop that will load them into a single data frame._

Vague question, asssuming binding by rows...

```{r, eval=FALSE}
results <- vector("list",length=length(files))
for(i in seq_along(files)) {
  results[i] <- read_csv(file=files[[i]])
}
results <- bind_rows(results)
```

### 2
_What happens if you use for (nm in names(x)) and x has no names? What if only some of the elements are named? What if the names are not unique?_

```{r}
x <- 1:10

for(nm in names(x)) {
  print(nm)
  print(x[[nm]])
}
```
blank!

```{r, error=TRUE}
names(x)[1:4] <- letters[1:4]
for(nm in names(x)) {
  print(nm)
  print(x[[nm]])
}
```

```{r, error=TRUE}
names(x)[5:6] <- c("e","e")
x
for(nm in names(x)) {
  print(x[[nm]])
}
```

multiple names return first element with that name

### 3
_Write a function that prints the mean of each numeric column in a data frame, along with its name. _

```{r, error=TRUE}
show_mean <- function(df) {
  for(i in seq_along(df)) {
    if(is.numeric(df[[i]])) {
      mean <- round(mean(df[[i]]),2)
      names(df)[[i]] %>% 
        str_c(":") %>% 
        str_pad(max(str_length(names(df))+2),"right") %>%
        str_c(mean) %>%
        print()
    }
  }
}

show_mean(iris)
```

### 4
_What does this code do? How does it work?_

```{r}
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}

mtcars
```

we have a list of functions.  The for loop goes through that list.  Because the function names matches the colnames of mtcars we can apply it specifically to those columns to transform them.

## 21.4.1

### 1.
_Read the documentation for apply(). In the 2d case, what two for loops does it generalise?_

doing something for every row or for every column of a matrix

### 2.
```{r}
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}

col_summary2 <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    if(is.numeric(df[[i]])) {
      out[i] <- fun(df[[i]])
    } else {
      out[i] <- NA
    }
  }
  out
}

col_summary2(iris,mean)

col_summary(iris,mean)
```



