---
title: "Apr 26 assignment"
author: "Julin N Maloof"
date: "4/25/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pre-requisites  
```{r}
library(tidyverse)
```

## 3.2

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### 3.2.4 Exercises

#### 1
```{r}
ggplot(mpg)
```

Generates a blank plot

#### 2
_how many rows and columns are in mtcar?_
```{r}
dim(mtcars)
```
32 rows and 11 columns

#### 3 
_what does the `drv` variable describe?_
```{r}
?mpg
```
`drv` indicates whether the car if front wheel, rear wheel, are four wheel drive

#### 4 
_make a scatterplot of `hwy` vs `cyl`_
```{r}
ggplot(mpg,aes(x=cyl,y=hwy)) + geom_point()
```

#### 5 
_what happens if you make a scatterplot of `class` vs `drv`?  Why is the plot not usefull
```{r}
ggplot(mpg,aes(x=class,y=drv)) + geom_point()
```

Not useful because we would want to know how many are in each combination.

### 3.3

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

### 3.3.1 exercises

#### 1
_Why are the points not blue?_
because we are manually setting the color but doing so inside the aes argument.  Instead:

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy),color="blue")
```

#### 2 
Which variable in `mpg` are categorical?  Which are continuous?
```{r}
junk <- sapply(colnames(mpg), function(x) {
  print(paste(x,
  if(is.numeric(get(x,mpg))) "continuous" else "categorical",
  sep=": "
  ))
})
```

####
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=cyl))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size=cyl))
```

Doesn't run:
```{r, eval=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape=cyl))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=displ,size=displ))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,stroke=displ))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=displ<5
                          ))
```

### 3.5.1

#### 1
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy)) + 
  facet_wrap(~displ)
```
A facet for every value

#### 2
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```
empty cells are combinations that do not exist

#### 3
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```


"." is a stand-in for nothing

#### 4

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

disadvantage: hard to see the relationship of the points. advantage: no overplotting

#### 5

nrow: number of rows, ncol: nomber of cols.  Can't set this for facet_grid because ther we go by the values.

#### 6 

screens are wider than they are tall