---
title: "June 28"
author: "Julin N Maloof"
date: "6/27/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)
library(magrittr)
```


# Chapter 13

## 13.3.1 Exercises

### 1

_Add a surrogate key to flights._

Because there are no unique combination of columns that can be a key we just create a sequential ID key
```{r}
flights <- flights %>% mutate(ID=row_number())
flights %>% select(ID, everything()) #so that we can put ID as the first column
```

### 2

_Identify the keys in the following datasets_

#### Lahman::Batting,

```{r}
?Lahman::Batting
Lahman::Batting %>% count(playerID,yearID,teamID) %>% filter(n>1) #nope
```

No!

What is the problem? Look at an example
```{r}
Lahman::Batting %>% filter(playerID=="anderjo01" & yearID==1898)
```
He played for two teams in the same season!

How about
```{r}
Lahman::Batting %>% count(playerID,yearID,stint) %>% filter(n>1)
```
Yes!

#### babynames::babynames

```{r}
library(babynames)
?babynames
head(babynames)
```

keys should be year, sex, and name

```{r}
babynames %>% count(year,sex,name) %>% arrange(desc(nn))
```

#### nasaweather::atmos

```{r}
library(nasaweather)
?atmos
head(atmos)
```

should be lat,long,year,month

```{r}
atmos %>% count(lat,long,year,month) %>% arrange(desc(n))
```

#### fueleconomy::vehicles

```{r}
library(fueleconomy)
head(vehicles)
```

```{r}
vehicles %>% count(id) %>% arrange(desc(n))
```

#### ggplot2::diamonds
```{r}
?diamonds
head(diamonds)
diamonds <- diamonds %>% mutate(ID=row_number()) %>% select(ID,everything())
diamonds
```


### 3

_Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers._

playerID connects `batting` to `Master` and connects `Salaries` to `master`

yearID,teamID,playerID connect `batting` to `Salaries`

Try this with the datamodelr package

```{r}
library(datamodelr)
library(Lahman)
```

Create data model
```{r}
dm_lahman <- dm_from_data_frames(Batting,Master,Salaries)
```

```{r}
dm_lahman %>% dm_create_graph() %>% dm_render_graph()
```

add references

```{r}
dm_lahman %<>% dm_add_references(Master$playerID==Batting$playerID,
                                 Batting$yearID==Salaries$yearID,
                                 Batting$playerID==Salaries$playerID,
                                 Batting$teamID==Salaries$teamID)
```

```{r}
dm_lahman %>% dm_create_graph(rankdir="LR") %>% dm_render_graph()
```

```{r}
dm_lahman %>% dm_create_graph(rankdir="LR",view_type="keys_only") %>% dm_render_graph()
```
Unfortunately the arrows do not go directly from key to key, but the "~" and underlines help sort it out...

```{r}
dm_lahman2 <- dm_from_data_frames(Master,Managers,AwardsManagers)
dm_lahman2 %>% dm_create_graph() %>% dm_render_graph()
```

set references
```{r}
dm_lahman2 %<>% dm_add_references(Managers$lgID==AwardsManagers$lgID,
                                  Managers$playerID==Master$playerID,
                                  Master$playerID==AwardsManagers$playerID,
                                  Managers$playerID==AwardsManagers$playerID,
                                  Managers$yearID==AwardsManagers$yearID)
```

```{r}
dm_lahman2 %>% dm_create_graph() %>% dm_render_graph()
```

```{r}
dm_lahman2 %>% dm_create_graph(view_type = "keys_only",rankdir="LR") %>% dm_render_graph()
```

Not perfect...can only have key going somewhere once, etc.

```{r}

```

## 13.4

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2
```

inner_join joins based on matching key and only retains those oberservations where the key matches.  __Unmatched rows are not included__

outer_joins keep additional unmatched observations.  left join keeps all observations in the left object, right join keeps all in the right object, and full join keeps both.

what about duplicate keys?

if one table has duplicate keys and the other doesn't then it is easy, the info from the non-duplicated table is added to each of the duplicated key table.

if both table has duplicate keys then you get all possible combinations. This is probably not what you want.

if you omit "by" then all column names present in both tables are used.

"by" can be a named character vector that specifys the relationships between the x and y columns

## 13.4.6 Exercises

### 1

_Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:_

```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

_You might want to use the size or colour of the points to display the average delay for each airport._


```{r}
flights %>% 
  group_by(dest) %>% 
  summarize(avg_delay = mean(arr_delay,na.rm=T)) %>%
  inner_join(airports, c("dest" = "faa")) %>%
  ggplot(aes(x=lon,y=lat,fill=avg_delay)) +
  borders("state") +
  geom_point(stroke=1,shape=21,size=3) +
  coord_quickmap() +
  scale_fill_gradient2(low="blue",mid="white",high="red") +
  xlim(c(-125,-65)) + ylim(c(25,50))
```

#### 2

_Add the location of the origin and destination (i.e. the lat and lon) to flights._

```{r}
airports %>% select(faa,lat,lon) %>%
  right_join(flights,by=c("faa" = "origin")) %>%
  rename(lat.origin=lat, lon.origin=lon) %>%
  left_join(airports, by=c("dest" = "faa")) %>%
  rename(lat.dest=lat, lon.dest=lon) %>%
  select(-faa, -name, -alt, starts_with("lat"), starts_with("lon"))
```

