---
title: "June 28"
author: "Julin N Maloof"
date: "6/27/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)
library(magrittr)
```


# Chapter 13

## 13.3.1 Exercises

### 1

_Add a surrogate key to flights._

Because there are no unique combination of columns that can be a key we just create a sequential ID key
```{r}
flights <- flights %>% mutate(ID=row_number())
flights %>% select(ID, everything()) #so that we can put ID as the first column
```

### 2

_Identify the keys in the following datasets_

#### Lahman::Batting,

```{r}
?Lahman::Batting
Lahman::Batting %>% count(playerID,yearID,teamID) %>% filter(n>1) #nope
```

No!

What is the problem? Look at an example
```{r}
Lahman::Batting %>% filter(playerID=="anderjo01" & yearID==1898)
```
He played for two teams in the same season!

How about
```{r}
Lahman::Batting %>% count(playerID,yearID,stint) %>% filter(n>1)
```
Yes!

#### babynames::babynames

```{r}
library(babynames)
?babynames
head(babynames)
```

keys should be year, sex, and name

```{r}
babynames %>% count(year,sex,name) %>% arrange(desc(nn))
```

#### nasaweather::atmos

```{r}
library(nasaweather)
?atmos
head(atmos)
```

should be lat,long,year,month

```{r}
atmos %>% count(lat,long,year,month) %>% arrange(desc(n))
```

#### fueleconomy::vehicles

```{r}
library(fueleconomy)
head(vehicles)
```

```{r}
vehicles %>% count(id) %>% arrange(desc(n))
```

#### ggplot2::diamonds
```{r}
?diamonds
head(diamonds)
diamonds <- diamonds %>% mutate(ID=row_number()) %>% select(ID,everything())
diamonds
```


### 3

_Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers._

playerID connects `batting` to `Master` and connects `Salaries` to `master`

yearID,teamID,playerID connect `batting` to `Salaries`

Try this with the datamodelr package

```{r}
library(datamodelr)
library(Lahman)
```

Create data model
```{r}
dm_lahman <- dm_from_data_frames(Batting,Master,Salaries)
```

```{r}
dm_lahman %>% dm_create_graph() %>% dm_render_graph()
```

add references

```{r}
dm_lahman %<>% dm_add_references(Master$playerID==Batting$playerID,
                                 Batting$yearID==Salaries$yearID,
                                 Batting$playerID==Salaries$playerID,
                                 Batting$teamID==Salaries$teamID)
```

```{r}
dm_lahman %>% dm_create_graph(rankdir="LR") %>% dm_render_graph()
```

```{r}
dm_lahman %>% dm_create_graph(rankdir="LR",view_type="keys_only") %>% dm_render_graph()
```
Unfortunately the arrows do not go directly from key to key, but the "~" and underlines help sort it out...

```{r}
dm_lahman2 <- dm_from_data_frames(Master,Managers,AwardsManagers)
dm_lahman2 %>% dm_create_graph() %>% dm_render_graph()
```

set references
```{r}
dm_lahman2 %<>% dm_add_references(Managers$lgID==AwardsManagers$lgID,
                                  Managers$playerID==Master$playerID,
                                  Master$playerID==AwardsManagers$playerID,
                                  Managers$playerID==AwardsManagers$playerID,
                                  Managers$yearID==AwardsManagers$yearID)
```

```{r}
dm_lahman2 %>% dm_create_graph() %>% dm_render_graph()
```

```{r}
dm_lahman2 %>% dm_create_graph(view_type = "keys_only",rankdir="LR") %>% dm_render_graph()
```

Not perfect...can only have key going somewhere once, etc.

```{r}

```

## 13.4

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2
```

inner_join joins based on matching key and only retains those oberservations where the key matches.  __Unmatched rows are not included__

outer_joins keep additional unmatched observations.  left join keeps all observations in the left object, right join keeps all in the right object, and full join keeps both.

what about duplicate keys?

if one table has duplicate keys and the other doesn't then it is easy, the info from the non-duplicated table is added to each of the duplicated key table.

if both table has duplicate keys then you get all possible combinations. This is probably not what you want.

if you omit "by" then all column names present in both tables are used.

"by" can be a named character vector that specifys the relationships between the x and y columns

## 13.4.6 Exercises

### 1

_Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:_

```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

_You might want to use the size or colour of the points to display the average delay for each airport._


```{r}
flights %>% 
  group_by(dest) %>% 
  summarize(avg_delay = mean(arr_delay,na.rm=T)) %>%
  inner_join(airports, c("dest" = "faa")) %>%
  ggplot(aes(x=lon,y=lat,fill=avg_delay)) +
  borders("state") +
  geom_point(stroke=1,shape=21,size=3) +
  coord_quickmap() +
  scale_fill_gradient2(low="blue",mid="white",high="red") +
  xlim(c(-125,-65)) + ylim(c(25,50))
```

#### 2

_Add the location of the origin and destination (i.e. the lat and lon) to flights._

```{r}
airports %>% select(faa,lat,lon) %>%
  right_join(flights,by=c("faa" = "origin")) %>%
  rename(lat.origin=lat, lon.origin=lon) %>%
  left_join(airports, by=c("dest" = "faa")) %>%
  rename(lat.dest=lat, lon.dest=lon) %>%
  select(-faa, -name, -alt, starts_with("lat"), starts_with("lon"))
```

### 3

_Is there a relationship between the age of a plane and its delays?_

```{r}
head(planes)
by_tail <- flights %>% group_by(tailnum) %>%
  summarize(avg_delay=mean(arr_delay,na.rm=TRUE))
planes %>% mutate(age=2013-year) %>% select(tailnum,age) %>%
  inner_join(by_tail,by="tailnum") %>%
  ggplot(aes(x=age,y=avg_delay)) + geom_point() + geom_smooth()
  
```

No relationship between age and delay

### 4

_What weather conditions make it more likely to see a delay?_

```{r}
head(weather)
head(flights)
```

Need to create an "hour" column in flights.  should be rounded to the nearest hour...

```{r}
flight_weather <- flights %>% mutate(sched_dep_hour=round(sched_dep_time%/%100 + # get the hour
                     (sched_dep_time %% 100) / 60,#get the minuted, convert to fraction
                     0)) %>% # round to 0 decimals 
  inner_join(weather,
             by=c("origin"="origin",
                  "year"="year",
                  "month"="month",
                  "day"="day",
                  "sched_dep_hour"="hour")) %>%
  select(ID, dep_delay, temp, dewp, humid, starts_with("wind"), precip, pressure, visib)
flight_weather
```

```{r}
flight_weather %>% select(-ID,dep_delay) %>%
  cor(flight_weather$dep_delay,use="pairwise.complete.obs")
```

```{r}
flight_weather %>% select(-ID,-dep_delay) %>%
  apply(2,function(x) {cor.test(x,flight_weather$dep_delay)$p.value})
```

everything is significant.  The largest correlations for delays are:

* high dewpoint (= fog)
* low pressure (= storm)
* low visibility

let's plot one of these

```{r}
qplot(visib,dep_delay,data=flight_weather,alpha=.1) + geom_smooth()
flight_weather %>% ggplot(aes(x=as.factor(visib),y=dep_delay)) + geom_boxplot()
```



## Exercises 13.5.1

### 1 
_What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)_

```{r}
flights %>% anti_join(planes,by = "tailnum")
flights %>% anti_join(planes,by = "tailnum") %>% 
  count(carrier)
```


### 2
_Filter flights to only show flights with planes that have flown at least 100 flights._

```{r}
flights %>% group_by(tailnum) %>% filter(n()>100)
```

### 3
_Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models._

```{r}
head(fueleconomy::common)
head(fueleconomy::vehicles)
fueleconomy::vehicles %>% semi_join(fueleconomy::common)
```

### 4
_Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?_

```{r}
library(RcppRoll)
flights %>% group_by(year,month,day,hour) %>% 
  summarize(mean_dep_delay_hour=mean(dep_delay,na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(mean_delay48=roll_mean(mean_dep_delay_hour,n=48,fill=NA)) %>%
  filter(mean_delay48==max(mean_delay48,na.rm=TRUE)) %>%
  semi_join(weather,.,by=c("month","day")) %>% summary() # but really we should look at this whole period...
```

### 5
_What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?_

```{r}
anti_join(flights, airports, by = c("dest" = "faa"))
```
These are flights where the destination airport is not listed in the airports table

```{r}
anti_join(airports, flights, by = c("faa" = "dest"))
```
These are airports that were not a destination for any flights.  (or not airports in many cases)

### 6
_You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above._

Is each plane flown by a single airline?  I don't see why I need a join function for this...
```{r}
flights %>% group_by(tailnum) %>% summarize(n_carriers = n_distinct(carrier)) %>% filter(n_carriers>1)
```
No: 18 planes were flown by two airlines

which airlines?
```{r}
flights %>% group_by(tailnum) %>% summarize(n_carriers = n_distinct(carrier)) %>% filter(n_carriers>1) %>%
  filter(!is.na(tailnum)) %>%
  semi_join(flights,.) %>%
  group_by(tailnum) %>%
  filter(!duplicated(carrier)) %>%
  select(tailnum, carrier) %>%
  left_join(airlines)

```

